---
title: "Functions-tutorial"
author: "RFH"
date: "February 13, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Creating functions
# Work with data in degrees F

- Convert to C

```{r}
airtemps <- c(212, 30.3, 78, 32, 0, -56)

test<-(airtemps-32)*5/9

celsius1 <- (airtemps[1]-32)*5/9 #how to do this for every value in the vector
celsius2 <- (airtemps[2]-32)*5/9
celsius3 <- (airtemps[3]-32)*5/9
```

Woah, that is way too much work!

# Write a function to do that
Use the 'function' function to write a function.

```{r}
fahr_to_celsius<- function(fahr) {
  celsius <- (fahr -32)* 5/9
  return(celsius) #return allows for a particularly value to be sent back
}
```


# Write a function to go C to F
```{r}
celsius_temps<- fahr_to_celsius(airtemps)
celsius_temps
```

Now test out how to do this in reverse and compare to the original values.

```{r}
celsius_to_fahr<- function(cels) {
  fahr <- (cels)*9/5+32
  return(fahr)
}

fahr_temps<- celsius_to_fahr(celsius_temps)
fahr_temps== airtemps #checks to make sure that the values are equal
```

# Show how to document functions

It is good practice to use ROxygen formatting to describe differnt functions

```{r}

#' Converts numeric temperature data from degrees Fahrenhiet to degrees Celsius
#'
#' @param fahr the value in Fahrenheit to be converted
#' @return the converted value in Celsius

fahr_to_celsius<- function(fahr) {
  celsius <- (fahr -32)* 5/9
  return(celsius) #return allows for a particularly value to be sent back
}
```

