---
title: "Practice Website"
author: "RFH"
date: "February 14, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning = FALSE, message=FALSE}
library(tidyverse)
library(scales)
```

# Reformat the data for plotting
It is good practice to all cite the data that you are reading in:

Alaska Department of Fish and Game, Division of Commercial Fisheries. 2017. Daily salmon escapement counts from the OceanAK database, Alaska, 1921-2017. Knowledge Network for Biocomplexity. doi:10.5063/F1S46Q6M.

```{r}
esc<- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e", method="libcurl"), stringsAsFactors = F )

```

* Calculate annual escapment by species and region
* Filter for just pacific salmon species (Sockeye, Chinook, Coho, Pink, Chum)

```{r, echo=FALSE}
annual_esc<-esc %>% 
  separate(sampleDate, into = c("Year", "Month", "Day"), sep = "-", remove = F) %>% 
  mutate(Year=as.numeric(Year)) %>% #coerce year to numeric typ
  group_by(Year, Species, SASAP.Region)%>% #split
  summarise(Count = sum(DailyCount)) %>% #apply and combine
  filter(Species %in% c("Sockeye","Chinook", "Coho", "Pink", "Chum"))

head(annual_esc)
```



# Generate some static plots

```{r, echo=FALSE}
ggplot(annual_esc, mapping=aes(x=Species, y= Count, fill=SASAP.Region))+
  geom_col()
```

```{r, echo=FALSE}
ggplot(filter(annual_esc, SASAP.Region =="Kodiak"), aes(x=Year, y=Count, color=Species))+
  geom_line()+
  geom_point()+
  scale_y_continuous(labels=comma)+
  ylab("Escapement (num of fish)")+
  xlab("Year")+
  ggtitle("Kodiak Salmon Escapement")+
  theme_test()
```

```{r, echo=FALSE}
p<-ggplot(annual_esc, aes(x=Year, y=Count, color=Species))+
  geom_line()+
  geom_point()+
  scale_y_continuous(label=comma)+
  facet_wrap(~SASAP.Region, scales = "free_y")+
  theme_test()

p

ggsave("Figures/regional_escapement.png", 
       plot=p,
       width= 7,
       height = 8,
       units = "in")

```

