---
title: "sf-package-training"
author: "RFH"
date: "February 15, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
library(sf)
library(tidyverse)
library(leaflet)
library(scales)
```

```{r}
ak_shp_sf<-read_sf("shapefiles/ak_regions_simp.shp")
```

Note: the sf package allows you to work with shapefiles as if they were dataframes, similar to an attribute table in ArcMap.

```{r}
plot(ak_shp_sf)
```

That looks like a weird projection. We can check this through different calls in the sf package.

```{r}
st_crs(ak_shp_sf) ##check projection and coordinate system

ak_shp_sf<-ak_shp_sf %>% 
  st_transform(crs = 3338)
```

#Spatial Joins!
```{r}
pop <- read.csv("shapefiles/alaska_population.csv")
```

Now we need to convert the .csv into an actual shapefile

```{r}
pop_sf<- st_as_sf(pop, 
                  coords = c("lng", "lat"),
                  crs=4326,
                  remove=F) %>% ##this retains the lat long columns
  st_transform(crs=3338)
```

st_join has options for how you would like to join shapefiles.

```{r}
pop_joined_sf<- st_join(pop_sf, ak_shp_sf, join=st_within)

plot(pop_joined_sf["region"])#this stops it from giving plots for every column
```

## Group by and summerize

```{r}
pop_region<-pop_joined_sf %>% 
  as.data.frame() %>% #drops geometry column
  group_by(region) %>% 
  summarize(total_pop = sum(population))

head(pop_region)
  
```

```{r}
ak_pop_sf <- left_join(ak_shp_sf, pop_region)

plot(ak_pop_sf["total_pop"])
```

Learning how to group by different categories

```{r}
ak_mgmt<- ak_pop_sf %>% 
  group_by(mgmt_area) %>% # here we do want to aggregate over geometries
  summarize(total_pop = sum(total_pop), do_union=F)

plot(ak_mgmt["total_pop"])
```


Write out shapefiles to working directory

```{r}
write_sf(ak_shp_sf, "shapefiles/ak_regions_populations.shp")
```

# Create static plots
```{r}
ggplot(ak_pop_sf)+
  geom_sf(data=ak_pop_sf, aes(fill=total_pop))+#put the aes in the geom_sf to help with layering
  geom_sf(data=pop_sf, size=0.5)+
  theme_bw()+
  labs(fill = "Total Population")+
  scale_fill_continuous(low= "khaki", high="firebrick", labels= comma)
```
## Interactive maps

Important note: you need to unproject your data and then tell leaflet how to project it.

The first step is to give leaflet a projection object that makes sense.

```{r leaflet functions}
epsg3338 <- leaflet::leafletCRS(
  crsClass = "L.Proj.CRS",
  code = "EPSG:3338",
  proj4def =  "+proj=aea +lat_1=55 +lat_2=65 +lat_0=50 +lon_0=-154 +x_0=0 +y_0=0 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs",
  resolutions = 2^(16:7))

```

```{r transform to remove projection}
ak_crs<- st_transform(ak_pop_sf, crs= 4326) #unprojected data with WGS84 datum

pop_crs<-st_transform(pop_sf, crs=4326)
```

```{r}
pal<- colorNumeric(palette = "Reds", domain= ak_crs$total_pop)

m<- leaflet(options = leafletOptions(crs=epsg3338)) %>% 
  addPolygons(data=ak_crs, 
              fillColor = ~pal(total_pop),
              weight=1,
              color="black",
              label=~region,
              fillOpacity = 1) %>% 
  addCircleMarkers(data= pop_crs,
                   lat=~lat,
                   lng=~lng,
                   radius = 2,
                   fillColor = "gray",
                   fillOpacity = 1,
                   weight = 0.25,
                   color="black",
                   label=~city)

m

```

